---
title: "Section 3: Applied Concepts"
author: "Gop Arop"
format: 
  html:
    embed-resources: true
---

## Class Prep
1. Hue is the simply the color, or the dominant wavelength. While intensity refers to the brightness of the color. 

2. One scenario where you'd want to use a sequential color scale is when plotting a heat map. A situation where a diverging color scale is suitable could be designing a visual where participants can only choose one of two outcomes. An example of an unordered color scale in use could be a visual representation of five different school majors, each assigned a distinct color.

3. In Figure 1.21, an instance of one gestalt inference taking precedence over another can be observed in the bottom right plot, where the lines joining the shapes create a perception of common fate, leading to the two shapes being perceived as a cohesive unit despite gaps in the line segments.

4. The results in Figure 1.23 suggest that charts leveraging length and position-based visual attributes demonstrate superior performance. This implies that the specific visual attributes of length and position contribute to the effectiveness of conveying information about a single categorical variable, favoring chart types that employ these characteristics.

5. I'd opt for using different colors instead of shapes. According to Figure 1.25, colors are more effective than shapes for distinguishing unordered categorical data, which could provide a clearer visual representation of the relationship across the three countries.

6. A time where the use of the left-side of Figure 1.27 would be helpful would be when trying to make an argument that significantly less people are going to law school. On the contrary, a time where the right-side of Figure 1.27 would also be helpful in showing that less people go to law school. However, there will be less of an emphasis on the rapid decline.

7. There are three main categories of problems with data visualization: aesthetic, substantive, and perceptual. Also, graphs that are ineffective can be due to poor aesthetics, poor data presentation , or perceptual issues. 
 
8. What are the key points that an individual engaged in everyday data visualization work should ensure before submitting their work?


## 3.2 Examples

# Example 1.
```{r}
library(tidyverse)
library(palmerpenguins)
theme_set(theme_minimal())

penguins_sum <- penguins |> group_by(species, year) |>
  summarise(n_penguins = n()) |>
  mutate(year = factor(year))

ggplot(data = penguins_sum, aes(x = year, fill = species)) +
  geom_col(aes(y = n_penguins)) +
  theme_minimal() +
  scale_fill_viridis_d()

ggplot(data = penguins_sum, aes(x = year, y = n_penguins,
                                colour = species, group = species)) +
  geom_line() +
  theme_minimal() +
  scale_colour_viridis_d()
```
1. Line plot overall more preferable because it's harder to see trends in the species of penguins that are not the Gentoo species (which is the baseline, not on a stable baseline).

# Example 2. 
```{r}
ggplot(data = penguins, aes(x = species)) +
  geom_bar(fill = "darkslategray4") +
  theme_minimal()

ggplot(data = penguins, aes(x = species)) +
  geom_bar(fill = "darkslategray4") +
  coord_cartesian(ylim = c(50, 160)) +
  theme_minimal()
```
2. The first plot is more preferable because, if our visualization encodes area of length, the y-axis should start at 0.

# Example 3.
```{r}
library(ggbeeswarm)
ggplot(data = penguins, aes(x = species, y = body_mass_g)) +
  geom_beeswarm(alpha = 0.7) +
  theme_minimal()

ggplot(data = penguins, aes(x = species, y = body_mass_g)) +
  geom_beeswarm(alpha = 0.7) +
  theme_minimal() +
  ylim(c(0, 6500))
```
3. First is more preferable because including 0 as a y limit for this type of graph gives alot of wasted space. 

# Example 4.
```{r}
ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point(aes(colour = species)) +
  scale_colour_viridis_d() +
  theme_minimal()

ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point() +
  facet_wrap( ~ species) +
  theme_minimal() 
```
4. The first is preferable because it better utilizes the position of points so we can easily see which species has high vs. low bill depths, etc. 

# Example 5.
```{r}
penguins <- penguins |> mutate(species_island_sex = interaction(species, 
                                                                island,
                                                                sex))
ggplot(data = penguins |>
         filter(!is.na(sex)), aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point(aes(colour = species_island_sex)) +
  scale_colour_viridis_d() +
  theme_minimal()


ggplot(data = penguins |> filter(!is.na(sex)), 
                                 aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point() +
  facet_wrap(~ species_island_sex) +
  theme_minimal()
```
5. The second one is more preferable because its hard for humans to accurately perceive 10 different color shades. 

# Revision of Example 5.
```{r}
ggplot(data = penguins |> filter(!is.na(sex)), aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point(aes(colour = sex, shape = species)) +
  facet_wrap(~island) +
  theme_minimal()
```

# Example 5.
The bar plot is stronger because it effectively uses position of the bars while the areas of the pie chart does not. 

# Example 6.
In the first, its easier to distinguish the region while the second has a higher data to ink ratio. 

## Your Turn 1

# Exercise 1.
a. You would use a sequential color scale
b. You would use a diverging color scale
c. You would use a qualitative color scale
d. You would use a diverging color scale

# Exercise 2.
```{r}
ggplot(data = penguins |> filter(!is.na(sex)), aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point(aes(colour = sex)) + scale_color_viridis_d() ## d is for discrete variables

ggplot(data = penguins |> filter(!is.na(sex)), aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point(aes(colour = body_mass_g)) + scale_color_viridis_c() ## c for continouous

ggplot(data = penguins |> filter(!is.na(sex)), aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point(aes(colour = body_mass_g)) + scale_color_viridis_b() ## b is for binning a continuous graph
```




